\vspace{-3em} \textbf{ Mapping stellar surfaces\\ II: An interpretable Gaussian process model for light curves }

### 2021-03-30 paper2/ms.tex
add line numbers

% !TeX root = ./ms.tex
\documentclass[modern]{aastex62}

@@

@@

\begin{abstract}
    The use of Gaussian processes (GPs) as models for astronomical time series datasets has recently
    become almost ubiquitous, given their ease of use and flexibility.
    GPs excel in particular at marginalization
    over the stellar signal in cases where the
    variability due to starspots rotating in and out of view is treated
    as a nuisance, such as in exoplanet transit modeling.
    However, these effective models are less useful in cases where the starspot signal is of primary interest since it is not obvious how the parameters of the GP model are related to the physical properties of interest, such
    as the size, contrast, and latitudinal distribution of the spots.
    Instead, it is common practice to explicitly model the effect
    of individual starspots on the light curve and attempt to infer their
    properties via optimization or posterior inference. Unfortunately,
    this process is
    degenerate, ill-posed, and often computationally intractable when
    applied to stars with more than a few spots and/or to ensembles of many
    light curves.
    %
    In this paper, we derive a closed-form expression for the
    mean and covariance of a Gaussian process model that describes
    the light curve of a rotating, evolving stellar surface
    conditioned on a given distribution of
    starspot sizes, contrasts, and latitudes.
    %
    We demonstrate that this model is correctly calibrated,
    allowing one to robustly infer physical parameters of interest from one
    or more stellar light curves, including
    the typical radii and the mean and variance of the latitude
    distribution of starspots.
    %
    Our GP has far-ranging implications for understanding the variability and
    magnetic activity of stars from both light curves and radial velocity (RV)
    measurements, as well as for robustly modeling correlated noise in both transiting
    and RV exoplanet searches.
    %
    Our implementation is efficient, user-friendly, and open source, available
    as the \Python package \starryprocess.
    \href{https://github.com/rodluger/starry_process}{\color{linkcolor}\faGithub}
\end{abstract}

@@

Despite whatever mystique the words ``Gaussian process'' may evoke, a GP
is nothing but a Gaussian distribution in many (formally infinite)
dimensions. Specifically, it is a Gaussian distribution over
\emph{functions} spanning a continuous domain (in our case, the time domain).
Similar to a multivariate Gaussian, which is described by
a $(K \times 1)$ vector $\pmb{\mu}$ characterizing
the mean of the process and a $(K \times K)$
matrix $\pmb{\Sigma}$ characterizing its covariance,
a GP is fully specified by a mean
function $m(t)$ and a kernel function $k(t, t')$.
To say that a random vector-valued variable $\mathbbb{f}$
defined on a $(K \times 1)$ time array $\mathbf{t}$
is ``distributed as a GP'' means that we may write
%
\begin{align}
    \label{eq:fnormal}
    \mathbbb{f} \sim \mathcal{N}\left( \pmb{\mu}, \pmb{\Sigma} \right)
    \quad,
\end{align}
%
where the elements of the mean and covariance are given by
$\mu_i = m(t_i)$ and $\Sigma_{i,j} = k(t_i, t_j)$, respectively.%
\footnote{%
    In this paper, we will use blackboard font (i.e., $\mathbbb{f}$) to
    denote random variables and serif font (i.e., $\mathbf{f}$) to denote
    particular realizations of those variables.
    See Appendix~\ref{sec:notation} for a detailed explanation of our notation.
}
Because of this relationship to multivariate Gaussians,
GPs are easy to sample from.%
\footnote{Given a 1-d array \texttt{mean} and a 2-d array \texttt{cov} in \Python,
    sampling from the corresponding GP (if it exists)
    can be done in a single line of code by calling
    \texttt{numpy.random.multivariate\_normal(mean, cov)}.}
But, as we alluded to earlier, the real showstopper is the application of GPs to inference problems.
Multivariate Gaussian distributions have a closed-form (marginal) likelihood
function, so it is easy to compute the probability
of one's data conditioned on a given value of $\pmb{\mu}$ and $\pmb{\Sigma}$
(i.e., the ``likelihood''; see Equation~\ref{eq:log-like} below).
This can in turn be maximized
to infer the optimal values of the model parameters
or used in a
numerical sampling scheme to compute the probability of those parameters
given the data (i.e., the ``posterior'').
Thanks to modern computer architectures, linear algebra packages, and
GP algorithms,
evaluating the GP likelihood may typically be done in a fraction of a second
for a reasonably-sized dataset (i.e., $K \lesssim 10^4$ datapoints).

Another big advantage of GPs is their flexibility. GPs are often dubbed
a class of ``non-parametric'' models, given that nowhere in the specification
of the GP is there an explicit functional form for $\mathbbb{f}$. Rather, a GP
is a stochastic
process whose draws can in principle take on \emph{any} functional form,
subject, however, to certain smoothness and correlation criteria
of tunable strictness
that are fully
encoded in the covariance $\pmb{\Sigma}$.
%
In many applications, particularly when modeling stellar light curves,
it is customary to restrict the problem by
assuming that the process is \emph{stationary}, such that we may write
%
\begin{align}
    \label{eq:kernel}
    \Sigma_{i,j} & = k(t_i, t_j)
    \nonumber                                    \\
                 & = k(\left| t_i - t_j \right|)
    \nonumber                                    \\
                 & \equiv k(\Delta t)
    \quad.
\end{align}
%
A stationary process is one that is independent of phase (or, in this case,
the actual value of the time $t$); rather, it depends only on the \emph{difference}
between the phases of two data points. The kernel of a stationary process is
therefore a one-dimensional function, typically chosen from a set of
standard functions with desirable smoothness and spectral properties.

@@

% Among a vast number of kernels are a few that are commonly used
% in modeling stellar variability: the exponential squared kernel,
% %
% \begin{align}
%     \label{eq:expsq}
%     k_\mathrm{E^2}(\Delta t)           & = \sigma^2 \exp\left(-\frac{\Delta t^2}{2\tau}\right)
%     \quad,                                                                                                                                           \\
%     \intertext{the Mat\'ern-3/2 kernel,}
%     \label{eq:mat32}
%     k_\mathrm{M^\frac{3}{2}}(\Delta t) & = \sigma^2 \left( 1 + \frac{\sqrt{3}\Delta t}{\tau} \right) \exp\left(-\frac{\sqrt{3}\Delta t}{\tau}\right)
%     \quad,                                                                                                                                           \\
%     \intertext{the exponential sine squared kernel,}
%     \label{eq:expsinsq}
%     k_\mathrm{S^2}(\Delta t)           & = \sigma^2 \exp\left[-\Gamma\sin^2\left( \frac{\pi \Delta t}{P} \right)\right]
%     \quad,                                                                                                                                           \\
%     \intertext{the cosine kernel,}
%     k_\mathrm{C}(\Delta t)             & = \sigma^2 \cos\left( \frac{2\pi \Delta t}{P} \right)
%     \quad,
% \end{align}
% %
% or any linear combination and/or product of these. In the expressions above,
% $\sigma$, $\tau$, $\Gamma$, and $P$ are
% the \emph{hyperparameters} of the kernels.
% Specifically, $\sigma^2$ is the variance (the value
% of the kernel when $t_i = t_j$), $\tau$ is a typical timescale over which
% data are correlated, $\Gamma$ is a dimensionless scale parameter, and $P$ is
% a period.
% Once a kernel has been chosen, these scalar quantities fully
% specify the covariance of the Gaussian process. GP-based inference
% entails optimizing or sampling over these few quantities, which in many
% cases can be done easily and efficiently.

@@

Before we dive into the computation of our GP, it is useful to introduce the spherical harmonics,
a set of orthogonal functions on the surface of the sphere which we will use
to describe the intensity field on the surface of a star
(Figure~\ref{fig:ylms}). As we will see below,
the spherical harmonics are a particularly convenient basis in which to
describe starspot distributions%
\footnote{There are, of course, drawbacks to using this basis:
    in particular, the spherical harmonics are smooth, continuous functions that
    struggle (at finite degree $l$) to capture high resolution features such as
    small starspots. We discuss this point at length in \S\ref{sec:tinyspots},
    where we show that our model is useful even when applied to stars with spots
    smaller than the effective resolution of the GP.}, as they will allow us to compute
moments of the intensity distribution analytically. Of more immediate concern,
\citet{Luger2019} showed that there is a linear relationship between the
spherical harmonic expansion of a stellar surface and the total disk-integrated
flux $\mathbf{f}$ (i.e., the light curve)
one would observe as the star rotates about a fixed axis.
If the stellar surface intensity is described by a spherical harmonic
coefficient vector $\mathbf{y}$ (up to a certain degree $l_\mathrm{max}$),
the flux is given by
%
\begin{align}
    \label{eq:fAy}
    \mathbf{f} = \mathbf{1} + \pmb{\mathcal{A}}(I, P, \mathbf{u}) \, \mathbf{y}
    \quad,
\end{align}
%
where $\mathbf{1}$ is the ones vector and
$\pmb{\mathcal{A}}$ is the \starry
design matrix, a purely linear operator that transforms from the spherical
harmonic basis to the flux basis; it is
a function of the stellar inclination $I$, the stellar
rotation period $P$, and the stellar limb darkening coefficients $\mathbf{u}$,
as well as the observation times (see Appendix~\ref{sec:starry} for details).

@@

\begin{align}

@@

\end{align}

@@

In this paper, we assume that the parameters $\pmb{\theta}_\bullet$
described above are the \emph{physically interesting} ones. That is,
given a light curve $\mathbf{f}$ or an ensemble of $M$ light curves of
statistically similar stars
$\left( \mathbf{f}_0 \,\, \mathbf{f}_1 \,\, \cdots \,\,  \mathbf{f}_{M-1} \right)^\top$,
we wish to infer the statistical properties of the starspots, encoded in
the entries of the vector $\pmb{\theta}_\bullet$.
%
This is typically a tall order, since it requires marginalizing over all
the nuisance parameters, which include the nitty-gritty details of the
size, contrast, and location of \emph{every spot} (and, if $M > 1$, on \emph{every star}
in the ensemble). Fortunately, however, the Gaussian process we constructed
does just that. Specifically, given the mean and covariance of the process,
we are able to directly evaluate the log marginal likelihood of the $m^\mathrm{th}$
dataset
conditioned on a specific value of $\pmb{\theta}_\bullet$ (as well as $I$,
$P$, and $\mathbf{u}$):
%
\begin{align}
    \label{eq:log-like}
    \ln \mathcal{L}_m\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
    =
     & -\frac{1}{2}
    \mathbf{r}_m^\top\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
    \big[
        \pmb{\Sigma}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
        +
        \mathbf{C}_m
        \big]^{-1}
    \mathbf{r}_m\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
    \nonumber       \\[0.75em]
     & -
    \frac{1}{2}
    \ln \Big|
    \pmb{\Sigma}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
    +
    \mathbf{C}_m
    \Big|
    -
    \frac{K}{2}
    \ln \left( 2 \pi \right)
    \quad,
\end{align}
%
where
%
\begin{align}
    \mathbf{r}_m\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
     & \equiv
    \mathbf{f}_m - \pmb{\mu}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
\end{align}
%
is the residual vector,
%
$\mathbf{C}_m$ is the data covariance
(which in most cases is a diagonal matrix whose entries
are the squared uncertainty corresponding to each data point in the light curve),
%
$| \cdots |$ denotes the determinant, and $K$ is the number of data points in
each light curve.%
\footnote{
    In Equation~(\ref{eq:log-like}) we implicitly assume all stars in the
    ensemble are observed at the same set of times $\mathbf{t}$.
    If this is not the case, the mean and covariance of the GP for each
    star must be computed from Equations~(\ref{eq:mean}) and (\ref{eq:cov})
    with the flux design matrix $\pmb{\mathcal{A}}(\mathbf{t}_m)$ evaluated
    at the particular observation times $\mathbf{t}_m$.
}
In an ensemble analysis, the joint marginal likelihood of all datasets is
simply the product of the individual likelihoods, so in log space we have
%
\begin{align}
    \ln \mathcal{L}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
     & =
    \sum_{m} \ln \mathcal{L}_m\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
    \quad.
\end{align}
%
The marginal likelihood may be interpreted as the probability of the data
given the model. Typically, we are interested in the reverse: the probability
of the \emph{model} given the \emph{data}, i.e., the posterior probability
distribution. In later sections we present a comprehensive suite of
posterior inference exercises demonstrating that our GP model is correctly
calibrated, allowing one to efficiently infer statistical properties of starspots
from light curves with minimal bias.

\subsection{Marginalizing over inclination}
\label{sec:gp-inc}
%
As we mentioned in the previous section,
the equations for the mean and covariance of our GP
(Equations~\ref{eq:mean_f} and \ref{eq:cov_f}, respectively) are conditioned
on specific values of the stellar and spot properties. To obtain the posterior distribution
for these parameters, we must typically resort to numerical sampling techniques,
which often scale steeply with the number of parameters. It is therefore generally
desirable to keep the total number of parameters small, especially when
employing the GP in an ensemble setting.
In such a setting, we might have light curves from $M$ stars, all of which
we believe to have similar spot properties (perhaps because they have
similar spectral types and rotation periods, for example).
The total number of parameters in our problem is therefore
%
\begin{align}
    N = 4 M + 5
    \quad,
\end{align}
%
since each of the stars will have their own set of 4 stellar properties
(an inclination, a period, and usually two limb darkening coefficients)
but will all share the same 5 spot properties
$\pmb{\theta}_\bullet$ (by assumption).
For a reasonably sized ensemble of $M=100$
stars, we would have to sample over $N = 405$ parameters.
While large, this number is certainly not absurd, especially by modern standards.
However, it does pose
a problem when considering how complex the posterior distribution for the
spot mapping problem can be. In addition to strong nonlinear degeneracies
between some of the parameters (such as the contrast $c$ and the
number of spots $n$), the posterior is often multimodal, especially in the
stellar inclinations. While modern sampling schemes such as Hamiltonian
Monte Carlo and Nested Sampling may in principle be able to deal with these
issues, in practice it can be very difficult to obtain convergence in a
reasonable amount of time.

@@

Fortunately, there is a better way to reduce the number of parameters in
the problem: we can explicitly marginalize over the stellar inclination.
That is, we may write the mean and covariance of our GP as
%
%
\begin{align}
    \label{eq:mu_marg}
    \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbbb{f} \, \Big| \, P, \mathbf{u}, \pmb{\theta}_\bullet \Big]
    \nonumber                                                                                                                                                                                                        \\
     & = \mathbf{1} + \mathbf{e}_I
    \\[0.5em]
    \label{eq:cov_marg}
    \pmb{\Sigma}(P, \mathbf{u}, \pmb{\theta}_\bullet)
     & = \mathrm{E} \Big[ \mathbbb{f} \, \mathbbb{f}^\top \, \Big| \, P, \mathbf{u}, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet) \pmb{\mu}^\top(P, \mathbf{u}, \pmb{\theta}_\bullet)
    \nonumber                                                                                                                                                                                                        \\
     & =
    \mathbf{E}_I
    -
    \mathbf{e}_I \,
    \mathbf{e}_I^\top
\end{align}
%
where we define the inclination first moment integral
%
\begin{align}
    \label{eq:eI}
    \mathbf{e}_I
     & \equiv
    \int
    \pmb{\mathcal{A}}(\mathbb{I}, P, \mathbf{u}) \,
    \mathrm{E} \Big[ \mathbbb{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
    p(\mathbb{I}) \, \mathrm{d}\mathbb{I}
\end{align}
%
and the inclination second moment integral
%
\begin{align}
    \label{eq:EI}
    \mathbf{E}_I
     & \equiv
    \int
    \pmb{\mathcal{A}}(\mathbb{I}, P, \mathbf{u}) \,
    \mathrm{E} \Big[ \mathbbb{y} \, \mathbbb{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] \,
    \pmb{\mathcal{A}}^\top(\mathbb{I}, P, \mathbf{u})
    p(\mathbb{I}) \, \mathrm{d}\mathbb{I}
    \quad,
\end{align}
%
and $\mathbb{I}$ is the random variable corresponding to the inclination.
%
The expectations inside the integrals in the expressions for
$\mathbf{e}_I$ and $\mathbf{E}_I$
are given by
Equations~(\ref{eq:exp_y}) and (\ref{eq:exp_yy}), respectively, and
are computed in Appendix~\ref{sec:integrals}.
%
If we are able to perform the integrals in those expressions,
we can \emph{dramatically} reduce the number of
parameters in our ensemble problem.
%
As we show in Appendix~\ref{sec:gp-inc}, if we assume that stellar
inclinations are distributed isotropically, these integrals
do in fact have closed-form solutions.

%
Finally, for future reference, it is useful to note that the mean of the GP
is constant:
%
\begin{align}
    \label{eq:scalar-mean}
    \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet)
     & =  (1 + e_I)\mathbf{1}
    \nonumber                 \\
     & \equiv
    \mu \, \mathbf{1}
    \quad,
\end{align}
%
since by construction our GP is longitudinally isotropic
(see Appendix~\ref{sec:inc-mom1}).

@@

\begin{proof}{test_ephi}

@@

\end{proof}

@@

\begin{proof}{test_eI}

>>>>

% !TeX root = ./ms.tex
\documentclass[modern,linenumbers]{aastex62}

@@

% Force line numbers
\usepackage{lineno}
\linenumbers

@@

\begin{abstract}
    The use of Gaussian processes (GPs) as models for astronomical time series datasets has recently
    become almost ubiquitous, given their ease of use and flexibility. GPs excel in particular at
    marginalization over the stellar signal when the variability due to starspots is treated as a
    nuisance, as in exoplanet transit modeling. However, these effective models are less useful in
    cases where the starspot signal is of primary interest, since it is not obvious how the parameters
    of the GP relate to physical parameters like the spot size, contrast, and latitudinal distribution.
    Instead, it is common practice to explicitly model the effect of individual starspots on the
    light curve and attempt to infer their properties via optimization or posterior inference.
    Unfortunately, this process is ill-posed and often computationally intractable when applied
    to stars with more than a few spots and/or to ensembles of many stars. Here we derive a
    closed-form expression for a GP that describes the light curve of a rotating, evolving stellar
    surface conditioned on a given distribution of starspot sizes, contrasts, and latitudes. We
    demonstrate that this model is correctly calibrated, allowing one to robustly infer physical
    parameters of interest from one or more light curves, including the typical spot radii and
    latitudes. Our GP has far-ranging implications for understanding the variability and magnetic
    activity of stars from light curves and radial velocity (RV) measurements, as well as for
    modeling correlated noise in exoplanet searches. Our implementation is efficient, user-friendly,
    and open source, available in the package \starryprocess.
    \href{https://github.com/rodluger/starry_process}{\color{linkcolor}\faGithub}
\end{abstract}

@@

Despite whatever mystique the words ``Gaussian process'' may evoke, a GP
is nothing but a Gaussian distribution in many (formally infinite)
dimensions. Specifically, it is a Gaussian distribution over
\emph{functions} spanning a continuous domain (in our case, the time domain).
Similar to a multivariate Gaussian, which is described by
a $(K \times 1)$ vector $\pmb{\mu}$ characterizing
the mean of the process and a $(K \times K)$
matrix $\pmb{\Sigma}$ characterizing its covariance,
a GP is fully specified by a mean
function $m(t)$ and a kernel function $k(t, t')$.
To say that a random vector-valued variable $\mathbbb{f}$
defined on a $(K \times 1)$ time array $\mathbf{t}$
is ``distributed as a GP'' means that we may write
%
\begin{linenomath}\begin{align}
        \label{eq:fnormal}
        \mathbbb{f} \sim \mathcal{N}\left( \pmb{\mu}, \pmb{\Sigma} \right)
        \quad,
    \end{align}\end{linenomath}
%
where the elements of the mean and covariance are given by
$\mu_i = m(t_i)$ and $\Sigma_{i,j} = k(t_i, t_j)$, respectively.%
\footnote{%
    In this paper, we will use blackboard font (i.e., $\mathbbb{f}$) to
    denote random variables and serif font (i.e., $\mathbf{f}$) to denote
    particular realizations of those variables.
    See Appendix~\ref{sec:notation} for a detailed explanation of our notation.
}
Because of this relationship to multivariate Gaussians,
GPs are easy to sample from.%
\footnote{Given a 1-d array \texttt{mean} and a 2-d array \texttt{cov} in \Python,
    sampling from the corresponding GP (if it exists)
    can be done in a single line of code by calling
    \texttt{numpy.random.multivariate\_normal(mean, cov)}.}
But, as we alluded to earlier, the real showstopper is the application of GPs to inference problems.
Multivariate Gaussian distributions have a closed-form (marginal) likelihood
function, so it is easy to compute the probability
of one's data conditioned on a given value of $\pmb{\mu}$ and $\pmb{\Sigma}$
(i.e., the ``likelihood''; see Equation~\ref{eq:log-like} below).
This can in turn be maximized
to infer the optimal values of the model parameters
or used in a
numerical sampling scheme to compute the probability of those parameters
given the data (i.e., the ``posterior'').
Thanks to modern computer architectures, linear algebra packages, and
GP algorithms,
evaluating the GP likelihood may typically be done in a fraction of a second
for a reasonably-sized dataset (i.e., $K \lesssim 10^4$ datapoints).

Another big advantage of GPs is their flexibility. GPs are often dubbed
a class of ``non-parametric'' models, given that nowhere in the specification
of the GP is there an explicit functional form for $\mathbbb{f}$. Rather, a GP
is a stochastic
process whose draws can in principle take on \emph{any} functional form,
subject, however, to certain smoothness and correlation criteria
of tunable strictness
that are fully
encoded in the covariance $\pmb{\Sigma}$.
%
In many applications, particularly when modeling stellar light curves,
it is customary to restrict the problem by
assuming that the process is \emph{stationary}, such that we may write
%
\begin{linenomath}\begin{align}
        \label{eq:kernel}
        \Sigma_{i,j} & = k(t_i, t_j)
        \nonumber                                    \\
                     & = k(\left| t_i - t_j \right|)
        \nonumber                                    \\
                     & \equiv k(\Delta t)
        \quad.
    \end{align}\end{linenomath}
%
A stationary process is one that is independent of phase (or, in this case,
the actual value of the time $t$); rather, it depends only on the \emph{difference}
between the phases of two data points. The kernel of a stationary process is
therefore a one-dimensional function, typically chosen from a set of
standard functions with desirable smoothness and spectral properties.

@@

% Among a vast number of kernels are a few that are commonly used
% in modeling stellar variability: the exponential squared kernel,
% %
% \begin{linenomath}\begin{align}
%     \label{eq:expsq}
%     k_\mathrm{E^2}(\Delta t)           & = \sigma^2 \exp\left(-\frac{\Delta t^2}{2\tau}\right)
%     \quad,                                                                                                                                           \\
%     \intertext{the Mat\'ern-3/2 kernel,}
%     \label{eq:mat32}
%     k_\mathrm{M^\frac{3}{2}}(\Delta t) & = \sigma^2 \left( 1 + \frac{\sqrt{3}\Delta t}{\tau} \right) \exp\left(-\frac{\sqrt{3}\Delta t}{\tau}\right)
%     \quad,                                                                                                                                           \\
%     \intertext{the exponential sine squared kernel,}
%     \label{eq:expsinsq}
%     k_\mathrm{S^2}(\Delta t)           & = \sigma^2 \exp\left[-\Gamma\sin^2\left( \frac{\pi \Delta t}{P} \right)\right]
%     \quad,                                                                                                                                           \\
%     \intertext{the cosine kernel,}
%     k_\mathrm{C}(\Delta t)             & = \sigma^2 \cos\left( \frac{2\pi \Delta t}{P} \right)
%     \quad,
% \end{align}\end{linenomath}
% %
% or any linear combination and/or product of these. In the expressions above,
% $\sigma$, $\tau$, $\Gamma$, and $P$ are
% the \emph{hyperparameters} of the kernels.
% Specifically, $\sigma^2$ is the variance (the value
% of the kernel when $t_i = t_j$), $\tau$ is a typical timescale over which
% data are correlated, $\Gamma$ is a dimensionless scale parameter, and $P$ is
% a period.
% Once a kernel has been chosen, these scalar quantities fully
% specify the covariance of the Gaussian process. GP-based inference
% entails optimizing or sampling over these few quantities, which in many
% cases can be done easily and efficiently.

@@

Before we dive into the computation of our GP, it is useful to introduce the spherical harmonics,
a set of orthogonal functions on the surface of the sphere which we will use
to describe the intensity field on the surface of a star
(Figure~\ref{fig:ylms}). As we will see below,
the spherical harmonics are a particularly convenient basis in which to
describe starspot distributions%
\footnote{There are, of course, drawbacks to using this basis:
    in particular, the spherical harmonics are smooth, continuous functions that
    struggle (at finite degree $l$) to capture high resolution features such as
    small starspots. We discuss this point at length in \S\ref{sec:tinyspots},
    where we show that our model is useful even when applied to stars with spots
    smaller than the effective resolution of the GP.}, as they will allow us to compute
moments of the intensity distribution analytically. Of more immediate concern,
\citet{Luger2019} showed that there is a linear relationship between the
spherical harmonic expansion of a stellar surface and the total disk-integrated
flux $\mathbf{f}$ (i.e., the light curve)
one would observe as the star rotates about a fixed axis.
If the stellar surface intensity is described by a spherical harmonic
coefficient vector $\mathbf{y}$ (up to a certain degree $l_\mathrm{max}$),
the flux is given by
%
\begin{linenomath}\begin{align}
        \label{eq:fAy}
        \mathbf{f} = \mathbf{1} + \pmb{\mathcal{A}}(I, P, \mathbf{u}) \, \mathbf{y}
        \quad,
    \end{align}\end{linenomath}
%
where $\mathbf{1}$ is the ones vector and
$\pmb{\mathcal{A}}$ is the \starry
design matrix, a purely linear operator that transforms from the spherical
harmonic basis to the flux basis; it is
a function of the stellar inclination $I$, the stellar
rotation period $P$, and the stellar limb darkening coefficients $\mathbf{u}$,
as well as the observation times (see Appendix~\ref{sec:starry} for details).

@@

\begin{linenomath}\begin{align}

@@

    \end{align}\end{linenomath}

@@

In this paper, we assume that the parameters $\pmb{\theta}_\bullet$
described above are the \emph{physically interesting} ones. That is,
given a light curve $\mathbf{f}$ or an ensemble of $M$ light curves of
statistically similar stars
$\left( \mathbf{f}_0 \,\, \mathbf{f}_1 \,\, \cdots \,\,  \mathbf{f}_{M-1} \right)^\top$,
we wish to infer the statistical properties of the starspots, encoded in
the entries of the vector $\pmb{\theta}_\bullet$.
%
This is typically a tall order, since it requires marginalizing over all
the nuisance parameters, which include the nitty-gritty details of the
size, contrast, and location of \emph{every spot} (and, if $M > 1$, on \emph{every star}
in the ensemble). Fortunately, however, the Gaussian process we constructed
does just that. Specifically, given the mean and covariance of the process,
we are able to directly evaluate the log marginal likelihood of the $m^\mathrm{th}$
dataset
conditioned on a specific value of $\pmb{\theta}_\bullet$ (as well as $I$,
$P$, and $\mathbf{u}$):
%
\begin{linenomath}\begin{align}
        \label{eq:log-like}
        \ln \mathcal{L}_m\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
        =
         & -\frac{1}{2}
        \mathbf{r}_m^\top\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
        \big[
            \pmb{\Sigma}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
            +
            \mathbf{C}_m
            \big]^{-1}
        \mathbf{r}_m\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
        \nonumber       \\[0.75em]
         & -
        \frac{1}{2}
        \ln \Big|
        \pmb{\Sigma}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
        +
        \mathbf{C}_m
        \Big|
        -
        \frac{K}{2}
        \ln \left( 2 \pi \right)
        \quad,
    \end{align}\end{linenomath}
%
where
%
\begin{linenomath}\begin{align}
        \mathbf{r}_m\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
         & \equiv
        \mathbf{f}_m - \pmb{\mu}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
    \end{align}\end{linenomath}
%
is the residual vector,
%
$\mathbf{C}_m$ is the data covariance
(which in most cases is a diagonal matrix whose entries
are the squared uncertainty corresponding to each data point in the light curve),
%
$| \cdots |$ denotes the determinant, and $K$ is the number of data points in
each light curve.%
\footnote{
    In Equation~(\ref{eq:log-like}) we implicitly assume all stars in the
    ensemble are observed at the same set of times $\mathbf{t}$.
    If this is not the case, the mean and covariance of the GP for each
    star must be computed from Equations~(\ref{eq:mean}) and (\ref{eq:cov})
    with the flux design matrix $\pmb{\mathcal{A}}(\mathbf{t}_m)$ evaluated
    at the particular observation times $\mathbf{t}_m$.
}
In an ensemble analysis, the joint marginal likelihood of all datasets is
simply the product of the individual likelihoods, so in log space we have
%
\begin{linenomath}\begin{align}
        \ln \mathcal{L}\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
         & =
        \sum_{m} \ln \mathcal{L}_m\left(I, P, \mathbf{u}, \pmb{\theta}_\bullet\right)
        \quad.
    \end{align}\end{linenomath}
%
The marginal likelihood may be interpreted as the probability of the data
given the model. Typically, we are interested in the reverse: the probability
of the \emph{model} given the \emph{data}, i.e., the posterior probability
distribution. In later sections we present a comprehensive suite of
posterior inference exercises demonstrating that our GP model is correctly
calibrated, allowing one to efficiently infer statistical properties of starspots
from light curves with minimal bias.

\subsection{Marginalizing over inclination}
\label{sec:gp-inc}
%
As we mentioned in the previous section,
the equations for the mean and covariance of our GP
(Equations~\ref{eq:mean_f} and \ref{eq:cov_f}, respectively) are conditioned
on specific values of the stellar and spot properties. To obtain the posterior distribution
for these parameters, we must typically resort to numerical sampling techniques,
which often scale steeply with the number of parameters. It is therefore generally
desirable to keep the total number of parameters small, especially when
employing the GP in an ensemble setting.
In such a setting, we might have light curves from $M$ stars, all of which
we believe to have similar spot properties (perhaps because they have
similar spectral types and rotation periods, for example).
The total number of parameters in our problem is therefore
%
\begin{linenomath}\begin{align}
        N = 4 M + 5
        \quad,
    \end{align}\end{linenomath}
%
since each of the stars will have their own set of 4 stellar properties
(an inclination, a period, and usually two limb darkening coefficients)
but will all share the same 5 spot properties
$\pmb{\theta}_\bullet$ (by assumption).
For a reasonably sized ensemble of $M=100$
stars, we would have to sample over $N = 405$ parameters.
While large, this number is certainly not absurd, especially by modern standards.
However, it does pose
a problem when considering how complex the posterior distribution for the
spot mapping problem can be. In addition to strong nonlinear degeneracies
between some of the parameters (such as the contrast $c$ and the
number of spots $n$), the posterior is often multimodal, especially in the
stellar inclinations. While modern sampling schemes such as Hamiltonian
Monte Carlo and Nested Sampling may in principle be able to deal with these
issues, in practice it can be very difficult to obtain convergence in a
reasonable amount of time.

@@

Fortunately, there is a better way to reduce the number of parameters in
the problem: we can explicitly marginalize over the stellar inclination.
That is, we may write the mean and covariance of our GP as
%
%
\begin{linenomath}\begin{align}
        \label{eq:mu_marg}
        \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet)
         & = \mathrm{E} \Big[ \mathbbb{f} \, \Big| \, P, \mathbf{u}, \pmb{\theta}_\bullet \Big]
        \nonumber                                                                                                                                                                                                        \\
         & = \mathbf{1} + \mathbf{e}_I
        \\[0.5em]
        \label{eq:cov_marg}
        \pmb{\Sigma}(P, \mathbf{u}, \pmb{\theta}_\bullet)
         & = \mathrm{E} \Big[ \mathbbb{f} \, \mathbbb{f}^\top \, \Big| \, P, \mathbf{u}, \pmb{\theta}_\bullet \Big] - \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet) \pmb{\mu}^\top(P, \mathbf{u}, \pmb{\theta}_\bullet)
        \nonumber                                                                                                                                                                                                        \\
         & =
        \mathbf{E}_I
        -
        \mathbf{e}_I \,
        \mathbf{e}_I^\top
    \end{align}\end{linenomath}
%
where we define the inclination first moment integral
%
\begin{linenomath}\begin{align}
        \label{eq:eI}
        \mathbf{e}_I
         & \equiv
        \int
        \pmb{\mathcal{A}}(\mathbb{I}, P, \mathbf{u}) \,
        \mathrm{E} \Big[ \mathbbb{y} \, \Big| \, \pmb{\theta}_\bullet \Big]
        p(\mathbb{I}) \, \mathrm{d}\mathbb{I}
    \end{align}\end{linenomath}
%
and the inclination second moment integral
%
\begin{linenomath}\begin{align}
        \label{eq:EI}
        \mathbf{E}_I
         & \equiv
        \int
        \pmb{\mathcal{A}}(\mathbb{I}, P, \mathbf{u}) \,
        \mathrm{E} \Big[ \mathbbb{y} \, \mathbbb{y}^\top \, \Big| \, \pmb{\theta}_\bullet \Big] \,
        \pmb{\mathcal{A}}^\top(\mathbb{I}, P, \mathbf{u})
        p(\mathbb{I}) \, \mathrm{d}\mathbb{I}
        \quad,
    \end{align}\end{linenomath}
%
and $\mathbb{I}$ is the random variable corresponding to the inclination.
%
The expectations inside the integrals in the expressions for
$\mathbf{e}_I$ and $\mathbf{E}_I$
are given by
Equations~(\ref{eq:exp_y}) and (\ref{eq:exp_yy}), respectively, and
are computed in Appendix~\ref{sec:integrals}.
%
If we are able to perform the integrals in those expressions,
we can \emph{dramatically} reduce the number of
parameters in our ensemble problem.
%
As we show in Appendix~\ref{sec:gp-inc}, if we assume that stellar
inclinations are distributed isotropically, these integrals
do in fact have closed-form solutions.

%
Finally, for future reference, it is useful to note that the mean of the GP
is constant:
%
\begin{linenomath}\begin{align}
        \label{eq:scalar-mean}
        \pmb{\mu}(P, \mathbf{u}, \pmb{\theta}_\bullet)
         & =  (1 + e_I)\mathbf{1}
        \nonumber                 \\
         & \equiv
        \mu \, \mathbf{1}
        \quad,
    \end{align}\end{linenomath}
%
since by construction our GP is longitudinally isotropic
(see Appendix~\ref{sec:inc-mom1}).

@@

\begin{linenomath}\begin{proof}{test_ephi}

@@

    \end{proof}\end{linenomath}

@@

\begin{linenomath}\begin{proof}{test_eI}
